\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath}
%\usepackage{color,hyperref}
\newcounter{address}
\setcounter{address}{1}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcommand{\eqnnumber}{equation}
\newcommand{\eg}{e.g.}
\newcommand{\ie}{i.e.}
\newcommand{\Ie}{I.e.}
\newcommand{\etal}{et~al.}
\newcommand{\Hipparcos}{\textit{Hipparcos}}
\newcommand{\Gaia}{\textit{Gaia}}

\begin{document}

\title{The velocity distribution of nearby stars from
  \Hipparcos\ data\\ III. The curious case of Arcturus}
\author{
  Jo~Bovy\altaffilmark{\ref{NYU},\ref{email}} \&
  David~W.~Hogg\altaffilmark{\ref{NYU},\ref{MPIA}}}
\altaffiltext{\theaddress}{\label{NYU}\stepcounter{address} Center for
  Cosmology and Particle Physics, Department of Physics, New York
  University, 4 Washington Place, New York, NY 10003}
\altaffiltext{\theaddress}{\stepcounter{address}\label{email} To whom
  correspondence should be addressed: \texttt{jo.bovy@nyu.edu}}
\altaffiltext{\theaddress}{\label{MPIA}\refstepcounter{address}
  Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117
  Heidelberg, Germany}

\begin{abstract}

\end{abstract}
\keywords{
Galaxy: abundances ---
Galaxy: kinematics and dynamics ---
Galaxy: structure ---
methods: statistical ---
Solar neighborhood ---
stars: kinematics
}


\section{Introduction}

Not much is know about the Arcturus group. Some history here.

We use our recent reconstruction of the velocity field to determine
probable Arcturus group members from a variety of data sets and look
at their properties.

\section{Data}

\section{Determing Arcturus membership}

\acknowledgments It is a pleasure to thank Sam Roweis for helpful
discussions and assistance.  Financial support for this project was
provided by the National Aeronautics and Space Administration (grant
NNX08AJ48G) and the National Science Foundation (grant
AST-0908357). DWH is a research fellow of the Alexander von
Humboldt Foundation of Germany.


%\bibliographystyle{apj}
%\bibliography{EM,bovy_streams,MML}


\end{document}
