\documentstyle[12pt,aaspp4]{article}

\begin{document}

\section{Observations}

\subsection{Early evidence for streams}

\cite{majewski94,majewski96} --- moving groups

\cite{irwin95,grillmair95} --- extra-tidal stars

\subsection{Sagittarius}

\cite{ibata94,ibata95,ibata97} --- discovery papers.

\cite{mateo96,alard96,alcock97,sdss1,martinez01} --- associated
RR Lyraes, now up to 40(?) degrees away.

\cite{mateo98,majewski99,sdss2} --- overdensities in color-magnitude space,
also up to 40 degrees away.

\cite{dohmp01} --- photometrically selected giants with follow-up
spectroscopy.

\cite{ibata01a} --- Carbon star stream from Sgr encircling the Galaxy.

\cite{jsh95,johnston99,helmi01,ibata01a,ibata01b} --- modelling.

\subsection{Ongoing and upcoming surveys}

\cite{sdss1,sdss2} --- Sloan

\cite{ggss1,spag} --- photometrically selected giants with follow
up spectroscopy: ``all-sky'' in equally-space bricks.

\cite{helmi00} --- prospects for FAME and GAIA.

\section{Theory}

\subsection{Numerical, analytic and semi-analytic descriptions of
evolution of tidal debris}

\cite{tremaine93} --- general estimates of scales involved.

\cite{jsh95} --- demonstration of persistence of moving
groups following tidal disruption.

\cite{jhb96} --- characteristics of debris seen in
simulations; rate of spreading as function of satellite
properties. 

\cite{johnston98a,johnston01c} --- application of Tremaine's work to
general spherical potentials and non-circular orbits; approximate
description based on energy and angular momentum properties 
only.

\cite{helmi99a} --- full linear-theory description in action-angle
variables.

\subsection{Applications}

\cite{jhb96} --- recovering streamers from catalogues of positions
of halo stars.

\cite{johnston99} --- intertpreting history of a satellite (Sgr)
using its tidal debris.

\cite{jsh99} --- measuring mass-loss-rates from satellites.

\cite{jzsh99} --- measuring the potential of the Milky Way using
tidal debris.

\cite{helmi99b} --- Finding streams in local kpc$^3$ using
HIPPAROS plus expectation that stream stars have similar
angular momenta.

\cite{helmi00} --- prospects for using FAME and GAIA to find 
streamers in action-space.

\cite{bullock01} --- is the degree of substructure seen in SDSS 
consistent with satellite destruction in CDM Universes?

\cite{johnston01c} --- how likely are we to find low-surface-brightness
streams around other galaxies within a CDM cosmological context?

%\bibliographystyle{apalike}
%\bibliography{tails,inprep,clusters,cosmo,dsph,misc,sgr,move}
\begin{thebibliography}{}

\bibitem[{Alard}, 1996]{alard96}
{Alard}, C. (1996).
\newblock Evidence for the sagittarius dwarf galaxy at low galactic latitudes.
\newblock {\em \apjl}, 458:L17--+.

\bibitem[{Alcock} et~al., 1997]{alcock97}
{Alcock}, C., {Allsman}, R.~A., {Alves}, D.~R., {Axelrod}, T.~S., {Becker},
  A.~C., {Bennett}, D.~P., {Cook}, K.~H., {Freeman}, K.~C., {Griest}, K.,
  {Guern}, J.~A., {Lehner}, M.~J., {Marshall}, S.~L., {Minniti}, D.,
  {Peterson}, B.~A., {Pratt}, M.~R., {Quinn}, P.~J., {Rodgers}, A.~W.,
  {Stubbs}, C.~W., {Sutherland}, W., {Welch}, D.~L., and {The MACHO
  Collaboration} (1997).
\newblock {MACHO Project Photometry of RR Lyrae Stars in the Sagittarius Dwarf
  Galaxy}.
\newblock {\em \apj}, 474:217+.

\bibitem[{Bullock} et~al., 2001]{bullock01}
{Bullock}, J.~S., {Kravtsov}, A.~V., and {Weinberg}, D.~H. (2001).
\newblock Hierarchical galaxy formation and substructure in the galaxy's
  stellar halo.
\newblock {\em \apj}, 548:33--46.

\bibitem[{Dohm-Palmer} et~al., 2001]{dohmp01}
{Dohm-Palmer}, R.~C., {Helmi}, A., {Morrison}, H., {Mateo}, M., {Olszewski},
  E.~W., {Harding}, P., {Freeman}, K.~C., {Norris}, J., and {Shectman}, S.~A.
  (2001).
\newblock {Mapping the Galactic Halo. V. Sagittarius Dwarf Spheroidal Tidal
  Debris 60\&deg; from the Main Body}.
\newblock {\em \apjl}, 555:L37--L40.

\bibitem[{Grillmair} et~al., 1995]{grillmair95}
{Grillmair}, C.~J., {Freeman}, K.~C., {Irwin}, M., and {Quinn}, P.~J. (1995).
\newblock Globular clusters with tidal tails: Deep two-color star counts.
\newblock {\em \aj}, 109:2553+.

\bibitem[{Helmi} and {de Zeeuw}, 2000]{helmi00}
{Helmi}, A. and {de Zeeuw}, P.~T. (2000).
\newblock Mapping the substructure in the galactic halo with the next
  generation of astrometric satellites.
\newblock {\em \mnras}, 319:657--665.

\bibitem[{Helmi} and {White}, 1999]{helmi99a}
{Helmi}, A. and {White}, S. D.~M. (1999).
\newblock Building up the stellar halo of the galaxy.
\newblock {\em \mnras}, 307:495--517.

\bibitem[{Helmi} and {White}, 2001]{helmi01}
{Helmi}, A. and {White}, S. D.~M. (2001).
\newblock Simple dynamical models of the sagittarius dwarf galaxy.
\newblock {\em \mnras}, 323:529--536.

\bibitem[{Helmi} et~al., 1999]{helmi99b}
{Helmi}, A., {White}, S. D.~M., {de Zeeuw}, P.~T., and {Zhao}, H. (1999).
\newblock Debris streams in the solar neighbourhood as relicts from the
  formation of the milky way.
\newblock {\em \nat}, 402:53--55.

\bibitem[{Ibata} et~al., 2001a]{ibata01b}
{Ibata}, R., {Irwin}, M., {Lewis}, G.~F., and {Stolte}, A. (2001a).
\newblock {Galactic Halo Substructure in the Sloan Digital Sky Survey: The
  Ancient Tidal Stream from the Sagittarius Dwarf Galaxy}.
\newblock {\em \apjl}, 547:L133--L136.

\bibitem[{Ibata} et~al., 2001b]{ibata01a}
{Ibata}, R., {Lewis}, G.~F., {Irwin}, M., {Totten}, E., and {Quinn}, T.
  (2001b).
\newblock Great circle tidal streams: Evidence for a nearly spherical massive
  dark halo around the milky way.
\newblock {\em \apj}, 551:294--311.

\bibitem[{Ibata} et~al., 1994]{ibata94}
{Ibata}, R.~A., {Gilmore}, G., and {Irwin}, M.~J. (1994).
\newblock A dwarf satellite galaxy in sagittarius.
\newblock {\em \nat}, 370:194+.

\bibitem[{Ibata} et~al., 1995]{ibata95}
{Ibata}, R.~A., {Gilmore}, G., and {Irwin}, M.~J. (1995).
\newblock Sagittarius: the nearest dwarf galaxy.
\newblock {\em \mnras}, 277:781--800.

\bibitem[{Ibata} et~al., 1997]{ibata97}
{Ibata}, R.~A., {Wyse}, R. F.~G., {Gilmore}, G., {Irwin}, M.~J., and
  {Suntzeff}, N.~B. (1997).
\newblock The kinematics, orbit, and survival of the sagittarius dwarf
  spheroidal galaxy.
\newblock {\em \aj}, 113:634--655.

\bibitem[{Irwin} and {Hatzidimitriou}, 1995]{irwin95}
{Irwin}, M. and {Hatzidimitriou}, D. (1995).
\newblock Structural parameters for the galactic dwarf spheroidals.
\newblock {\em \mnras}, 277:1354--1378.

\bibitem[{Ivezi{\' c}} et~al., 2000]{sdss1}
{Ivezi{\' c}}, {\v Z}., {Goldston}, J., {Finlator}, K., {Knapp}, G.~R.,
  {Yanny}, B., {McKay}, T.~A., {Amrose}, S., {Krisciunas}, K., {Willman}, B.,
  {Anderson}, S., {Schaber}, C., {Erb}, D., {Logan}, C., {Stubbs}, C., {Chen},
  B., {Neilsen}, E., {Uomoto}, A., {Pier}, J.~R., {Fan}, X., {Gunn}, J.~E.,
  {Lupton}, R.~H., {Rockosi}, C.~M., {Schlegel}, D., {Strauss}, M.~A., {Annis},
  J., {Brinkmann}, J., {Csabai}, I.~., {Doi}, M., {Fukugita}, M., {Hennessy},
  G.~S., {Hindsley}, R.~B., {Margon}, B., {Munn}, J.~A., {Newberg}, H.~J.,
  {Schneider}, D.~P., {Smith}, J.~A., {Szokoly}, G.~P., {Thakar}, A.~R.,
  {Vogeley}, M.~S., {Waddell}, P., {Yasuda}, N., and {York}, D.~G. (2000).
\newblock Candidate rr lyrae stars found in sloan digital sky survey
  commissioning data.
\newblock {\em \aj}, 120:963--977.

\bibitem[{Johnston}, 1998]{johnston98a}
{Johnston}, K.~V. (1998).
\newblock A prescription for building the milky way's halo from disrupted
  satellites.
\newblock {\em \apj}, 495:297+.

\bibitem[{Johnston} et~al., 1996]{jhb96}
{Johnston}, K.~V., {Hernquist}, L., and {Bolte}, M. (1996).
\newblock Fossil signatures of ancient accretion events in the halo.
\newblock {\em \apj}, 465:278+.

\bibitem[{Johnston} et~al., 1999a]{johnston99}
{Johnston}, K.~V., {Majewski}, S.~R., {Siegel}, M.~H., {Reid}, I.~N., and
  {Kunkel}, W.~E. (1999a).
\newblock Constraining the history of the sagittarius dwarf galaxy using
  observations of its tidal debris.
\newblock {\em \aj}, 118:1719--1726.

\bibitem[{Johnston} et~al., 2001]{johnston01c}
{Johnston}, K.~V., {Sackett}, P.~D., and {Bullock}, J.~S. (2001).
\newblock Interpreting debris from satellite disruption in external galaxies.
\newblock {\em \apj, in press}.

\bibitem[{Johnston} et~al., 1999b]{jsh99}
{Johnston}, K.~V., {Sigurdsson}, S., and {Hernquist}, L. (1999b).
\newblock Measuring mass-loss rates from galactic satellites.
\newblock {\em \mnras}, 302:771--789.

\bibitem[{Johnston} et~al., 1995]{jsh95}
{Johnston}, K.~V., {Spergel}, D.~N., and {Hernquist}, L. (1995).
\newblock The disruption of the sagittarius dwarf galaxy.
\newblock {\em \apj}, 451:598+.

\bibitem[{Johnston} et~al., 1999c]{jzsh99}
{Johnston}, K.~V., {Zhao}, H., {Spergel}, D.~N., and {Hernquist}, L. (1999c).
\newblock Tidal streams as probes of the galactic potential.
\newblock {\em \apjl}, 512:L109--L112.

\bibitem[{Majewski} et~al., 1994]{majewski94}
{Majewski}, S.~R., {Munn}, J.~A., and {Hawley}, S.~L. (1994).
\newblock {Absolute proper motions to B approximately 22.5: Evidence for
  kimematical substructure in halo field stars}.
\newblock {\em \apjl}, 427:L37--L41.

\bibitem[{Majewski} et~al., 1996]{majewski96}
{Majewski}, S.~R., {Munn}, J.~A., and {Hawley}, S.~L. (1996).
\newblock {Absolute Proper Motions to B approximately 22.5: Large-Scale
  Streaming Motions and the Structure and Origin of the Galactic Halo}.
\newblock {\em \apjl}, 459:L73--+.

\bibitem[{Majewski} et~al., 2000]{ggss1}
{Majewski}, S.~R., {Ostheimer}, J.~C., {Kunkel}, W.~E., and {Patterson}, R.~J.
  (2000).
\newblock Exploring halo substructure with giant stars. i. survey description
  and calibration of the photometric search technique.
\newblock {\em \aj}, 120:2550--2568.

\bibitem[{Majewski} et~al., 1999]{majewski99}
{Majewski}, S.~R., {Siegel}, M.~H., {Kunkel}, W.~E., {Reid}, I.~N., {Johnston},
  K.~V., {Thompson}, I.~B., {Landolt}, A.~U., and {Palma}, C. (1999).
\newblock Starcounts redivivus. iii. a possible detection of the sagittarius
  dwarf spheroidal galaxy at b=-40 deg.
\newblock {\em \aj}, 118:1709--1718.

\bibitem[{Mart{\' i}nez-Delgado} et~al., 2001]{martinez01}
{Mart{\' i}nez-Delgado}, D., {Aparicio}, A., {G{\' o}mez-Flechoso}, M.~{\' A}.,
  and {Carrera}, R. (2001).
\newblock {Tidal Streams in the Galactic Halo: Evidence for the Sagittarius
  Northern Stream or Traces of a New Nearby Dwarf Galaxy}.
\newblock {\em \apjl}, 549:L199--L202.

\bibitem[{Mateo} et~al., 1996]{mateo96}
{Mateo}, M., {Mirabal}, N., {Udalski}, A., {Szymanski}, M., {Kaluzny}, J.,
  {Kubiak}, M., {Krzeminski}, W., and {Stanek}, K.~Z. (1996).
\newblock Discovery of a tidal extension of the sagittarius dwarf spheroidal
  galaxy.
\newblock {\em \apjl}, 458:L13--+.

\bibitem[{Mateo} et~al., 1998]{mateo98}
{Mateo}, M., {Olszewski}, E.~W., and {Morrison}, H.~L. (1998).
\newblock {Tracing the Outer Structure of the Sagittarius Dwarf
  Galaxy:Detections at Angular Distances between 10 deg and 34 deg}.
\newblock {\em \apjl}, 508:L55--L59.

\bibitem[{Morrison} et~al., 2000]{spag}
{Morrison}, H.~L., {Mateo}, M., {Olszewski}, E.~W., {Harding}, P.,
  {Dohm-Palmer}, R.~C., {Freeman}, K.~C., {Norris}, J.~E., and {Morita}, M.
  (2000).
\newblock Mapping the galactic halo. i. the ``spaghetti'' survey.
\newblock {\em \aj}, 119:2254--2273.

\bibitem[{Tremaine}, 1993]{tremaine93}
{Tremaine}, S. (1993).
\newblock A pessimist's view of galactic structure.
\newblock In {\em AIP Conf. Proc. 278: Back to the Galaxy}, pages 599+.

\bibitem[{Yanny} et~al., 2000]{sdss2}
{Yanny}, B., {Newberg}, H.~J., {Kent}, S., {Laurent-Muehleisen}, S.~A., {Pier},
  J.~R., {Richards}, G.~T., {Stoughton}, C., {Anderson}, J.~E., {Annis}, J.,
  {Brinkmann}, J., {Chen}, B., {Csabai}, I.~., {Doi}, M., {Fukugita}, M.,
  {Hennessy}, G.~S., {Ivezi{\' c}}, {\v Z}., {Knapp}, G.~R., {Lupton}, R.,
  {Munn}, J.~A., {Nash}, T., {Rockosi}, C.~M., {Schneider}, D.~P., {Smith},
  J.~A., and {York}, D.~G. (2000).
\newblock Identification of a-colored stars and structure in the halo of the
  milky way from sloan digital sky survey commissioning data.
\newblock {\em \apj}, 540:825--841.

\end{thebibliography}


\end{document}







